\begin{itemize}[leftmargin=-.5in]
\item[] General Background
    \subitem \textbf{E.A. Lee.} The Problem with Threads. doi:10.1109/MC.2006.180
    \subitem \textbf{Bonomi \& al.} Fog computing and its role in the internet of things. doi:10.1145/2342509.2342513
    \subitem \textbf{D. Perino \& al.} a content router for high-speed forwarding on content names
    \subitem \textbf{Zhou \& al.} Scheduling of Parallelized Synchronous Dataflow Actors
\item[] Bachelor's theses
    \subitem \textbf{Teemaa, Taavi} Rinnakkaisuuden mallit ohjelmointikielissä
    \subitem \textbf{Kiljunen, Olli} Tehtävärinnakkaisuus ohjelmoinnissa
\item[] Master's theses
    \subitem \textbf{Hanhirova, Jussi} Performance analysis of hardware accelerated scheduling
    \subitem \textbf{Saarinen, Joonas} Parallel Processing of Vehicle Telemetric Data
    \subitem \textbf{Rasa, Marko} Instrumentation of OpenMP task scheduling
\item[] Texas Instruments Documentation
    \subitem \textbf{Texas Instruments} SPRUGR9H KeyStone Architecture Multicore Navigator User's Guide
    \subitem \textbf{Texas Instruments} SPRUGW0C TMS320C66x DSP CorePac User Guide
    \subitem \textbf{Texas Instruments} SPRUGW7A KeyStone Architecture Multicore Shared Memory Controller (MSMC)
    \subitem \textbf{Texas Instruments} SPRUGZ2A KeyStone I Architecture Debug and Trace User Guide
    \subitem \textbf{Texas Instruments} TMS320C6678 datasheet
\item[] Dataflow \& Actors
    \subitem \textbf{Lee, Seshia} Introduction to Embedded Systems
\item[] Preesm
    \subitem \textbf{Pelcat M. et. al.} Preesm: A dataflow-based rapid prototyping framework for simplifying multicore DSP programming 10.1109/EDERC.2014.6924354
\item[] OpenEM
    \subitem \textbf{Texas Instruments} Open Event Machine Library API Spec
    \subitem \textbf{Texas Instruments} Open Event Machine library User Guide
    \subitem \textbf{Texas Instruments} OpenEM White Paper
    \subitem \textbf{F. Moerman.} Open event machine: A multi-core run-time designed for performance
    \subitem \textbf{Stotzer \& al.} OpenMP on the Low-Power TI Keystone II ARM/DSP System-on-Chip
\item[] Static and Dynamic Analysis
    \subitem \textbf{Wilhelm et al.} The Worst-Case Execution-Time Problem—Overview of Methods and Survey of Tools
    \subitem \textbf{Banks et al.} Introduction to Discrete-Event Simulation
    \subitem \textbf{Gustavsson et al.} Timing Analysis of Parallel Software Using Abstract Execution
    \subitem \textbf{Fujimoto et al.} Parallel Discrete-Event Simulation
\end{itemize}
