The filter applications process two video streams simultaneously as described in chapter \ref{chapter:construction}. One video stream is processed with a sobel filter and the other is processed with a gaussian filter. The dynamic behavior of the applications is investigated using different workloads.

\subsubsection{First Experiment}
\label{subsubsec:first-experiment}
The workloads used in the first expriment are presented in the table \ref{tab:preesm_setups}. The purpose of the different bitrates used for each video stream is to expose the behavior of the OpenEM scheduler in handling dynamic workloads. The static schedule in the PREESM application will provide a baseline to reflect the OpenEM performance to, but the performance of the applications should not be directly compared due to the difference in the runtime systems.

\begin{table}
    \begin{center}
        \begin{tabular}{ c c }
            Sobel Resolution & Gauss Resolution \\ \hline
            CIF              & CIF              \\ \hline
            4CIF             & CIF              \\ \hline
            CIF              & 4CIF             \\ \hline
            QCIF             & QCIF             \\ \hline
        \end{tabular}
        \caption{PREESM and OpenEM measurement setups}
        \label{tab:preesm_setups}
    \end{center}
\end{table}

\subsubsection{Second Experiment}
\label{subsubsec:second-experiment}
In the second experiment the OpenEM application is measured with different core masks to investigate the ability of the dynamic scheduler to utilize the increased parallel resources. Both filters of the OpenEM application are loaded with CIF streams. Different numbers of cores are used, but all actors can run on all available cores. The experiment is run with core masks allowing one to eight cores be used for processing the streams.

\subsubsection{Third Experiment}
\label{subsubsec:third-experiment}
To examine the capabilities of the dynamic scheduler the the number of simultaneously processed frames is varied in the third experiment. The maximum number of frames that can be processed simultaneously corresponds to the number of initial events created in the initialization. Both filters are loaded with CIF streams and the measurements are run with 1 to 24 initial events. The number of simultaneous frames changes the throughput versus latency balance of the OpenEM application.
